\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{html}

\begin{document}

% $Header$

\title{The Colossus Battlelands Builder}

\author{Romain Dolbeau}

\maketitle

\section{BattlelandsBuilder}

BattlelandsBuilder is a (crude) tool to interactively
create Battlelands for use with the
``\htmladdnormallink{Colossus}{http://colossus.sf.net/}'' Game.
It has the same requirements as Colossus itself, see
\htmladdnormallink{Colossus README}{../../README/index.html}
and
\htmladdnormallink{Colossus build-HOWTO}{../../build/index.html}.

To create the program, just type ``make tools'' or ``and tools''.
See the
\htmladdnormallink{Colossus build-HOWTO}{../../build/index.html}
for more details.
This should create all the tools for Colossus, including
BattlelandsBuilder.

You can use it directly, starting from a bare plain,
or you can give a pre-existing Battlelands as a
starting point. Just give the path to the file as
the first argument to the command.

Exemples:
\texttt{java -jar BattlelandsBuilder.jar}
\texttt{java -jar BattlelandsBuilder.jar Battlelands/Woods}

Editing is point-and-click : click in the middle
of an hex to get a popup-menu to change the type
and elevation of the hex ; click near the border of
an hex to get a popup-menu to change this border of
this hex (warning: beware slope or wall who can
easily be put on the wrong hex, as there's not much
visual clue of their orientation).

\begin{itemize}

\item The ``Load Map'' menu item is used to load a Battlelands 
file in the map. The loaded battleland is \emph{merged}
with the currently displayed battleland. Use ``Erase Map''
below if you want to start from scratch.

\item The ``Save As'' menu item save the Battleland to a
file choosen by the user.

\item The ``Show Battleland'' menu item dump the Battleland
on standard output.

\item The ``Erase Map'' menu item return the map to the
default, i.e. only Plains of elevation 0, and no
hexside.

\item The ``Randomize Map'' menu item is used to generate
a random map from a description file. See the
\htmladdnormallink{BattlelandsRandomizer}{../BattlelandsRandomizer/index.html}
for a description of the input file.

\end{itemize}

\section{Frequently Asked Questions}

\begin{enumerate}

\item I can't get the Hex Side popup-menu !
\item My Hex Side is on the wrong hex !

To get the Hex Side menu, you must be quite close
to the Side itself. You must also not be on one
of the outer side, without a neighbor Hex, as you
can't put an Hex Side there. Finally, if the Hex
Side is put ``backward'', that means you clicked on
the wrong Hex ; for instance, you must click behind
the straight border, on the Sand, to get a proper
Dune. Clicking where you intend to get the round
part will put the Dune backward.

\item BattlelandsBuilder let me do dumb thing !

Well, yes. It's designed that way : you can do
whatever you please with it, including surrounding
a Tree with dune or filling the map with Volcanoes.
It will try to flag some obviously ``dangerous''
or ``wrong'' combinations with warnings, but you
can use the output anyway (the warnings are
comments that will be ignored by Colossus).

\end{enumerate}

\end{document}
