\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{html}

\begin{document}

% $Header$

\title{Hardware and Software requirements for Colossus}

\author{Romain Dolbeau, David Ripton}

\maketitle

\section{Requirements}

This document covers the basic hardware and software requirements to play Colossus. If you want to rebuild the game, please check the document ``\htmladdnormallink{Building Colossus}{../build/index.html}''.

Colossus should work with any Java Runtime Environment (\texttt{JRE}) or Java Development Kit (\texttt{JDK}) release 1.3 or newer. Using 1.4 is recommended, in particular on the Linux operating system, as it fixes some bugs. ``\htmladdnormallink{Building Colossus}{../build/index.html}'' contains the links to the various \texttt{JDK} and \texttt{JRE} available.

On the hardware side, you'll need a color display (the higher the resolution the better, below 1024x769 is likely to look ugly) and enough RAM (64 MibiBytes \cite{NIST-RCUU} may work, 128 MiB or more is better). A fast processor makes the game more comfortable, but as Colossus is turn-base, not real-time, it should work on about anything. For x86-family (i.e. Intel Pentium, AMD Athlon, etc.) processor, a speed of  at least 200 MegaHertz should allow the game to run decently.

Colossus is free, so in doubt, you can always try and see for yourself.

As a reference, you can see the table \ref{TESTED} for configurations that have already been tested. If you test another configuration, feel free to report it to the Colossus-Developer mailing list (see \htmladdnormallink{http://sourceforge.net/mail/?group\_id=1939}{http://sourceforge.net/mail/?group\_id=1939}, or the link from the Colossus Homepage at \htmladdnormallink{http://colossus.sourceforge.net/}{http://colossus.sourceforge.net/}).

\begin{table}
\begin{tabular}{|p{2.5cm}|p{2.5cm}|p{1.8cm}|p{2.5cm}|p{2.5cm}|p{1.8cm}|p{1.8cm}|}

\hline

\textbf{Operating System} & \textbf{CPU} & \textbf{memory} &  \textbf{\texttt{JRE}/\texttt{JDK} version} & \textbf{Colossus version} & \textbf{Comments}\\

\hline

MacOS X 10.1.3 & 667 Mhz PPC 7450 (G4) & 512 MiB & Apple-supplied 1.3.1 & 10/03/2002 & Playable \\

\hline

Linux/PPC, kernel 2.4.18 & 192 Mhz PPC 604e & 128 MiB & Blackdown 1.3.1 PPC & 10/03/2002 & Slow \\

\hline

Solaris 8 & 200 Mhz UltraSPARC I & 384 MiB & Sun 1.3.1 & 10/03/2002 & Playable \\

\hline

Solaris 7 & 440 Mhz UltraSPARC II & 512 MiB & Sun 1.4.0 & 10/03/2002 & Smooth \\

\hline

\end{tabular}
\caption{Some already tested configuration}
\label{TESTED}
\end{table}

\addcontentsline{toc}{section}{References}
\bibliographystyle{plain}
\bibliography{HW-SW-Req}

\end{document}
